---
title: "Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
learnr::run_tutorial
knitr::opts_chunk$set(echo = FALSE)
```


## Data Wrangling
Data wrangling is the process of transforming and mapping data from one “raw” data form into another format with the intent of making it more appropriate and valuable for a variety of downstream purposes such as analytics. Data wrangling is very useful in coding, especially when dealing with a large data set. To display how one uses R functions to wrangle data, we will be using the Fandango data set from the FiveThirtyEight package. The Fandango data set is a data frame with 146 rows representing movies and 23 variables. These 23 variables range from the film name, year, and various ratings. The file will take the reader on a step by step tutorial on how to data wrangle.

### Packages needed

packages that will be needed (run code below)

```{r packages, exercise=TRUE}
library(tidyverse) # Automatically loads ggplot2 and dplyr
library(fivethirtyeight)
```

### Terms of Data Wrangling 

dplyr is an excellent package for data wrangling. It is simple for helping to understand and work through manipulating data. One of the most significant advantages of this package is the pipe function. You can use the pipe function “%>%” to write a pipeline to combine different functions in R. You can even save “piped” operations as a new data frame.

I know it can be confusing all of the words used with the pipe operation in dplyr, so to make things easier, dplyr provides a grammar for data transformation. 

Five verbs for working with data in a data frame:
- select(): take (selects) a subset of columns (variables) 
- filter(): take a subset of rows (observations, cases, individuals), a way to filter the data to only see what you are looking for. 
- mutate(): add or modify existing columns (can even use this function to compute new variables through a provided equation) 
- arrange(): sort rows (descending or ascending order, etc.) 
- summarize(): aggregate data across rows (e.g., grouped means)
### Select
*First, we will try to use the select function. Say you just wanted to know the name of the movie and the year of the film, and its rotten tomato score. You would write the pipline as below:* 

Run Code
```{r select, exercise=TRUE, exercise.lines = 5}
movie_year = fandango %>%
  select(film, year, rottentomatoes)
movie_year
```
It displays all 146 movies the year or the film and the rotten tomato score, pretty cool, right? But 146 rows is still a lot of data.

### Filter

*Next, the filter function allows you to take the data that you selected and filter it to take a subset of rows to only see what you are looking for. The filter function can be used alone or in conjunction with the select function*

Run Code
```{r Filter, exercise=TRUE, exercise.lines = 5}
below_fifty = fandango %>%
  select(film, year, rottentomatoes) %>%
  filter(rottentomatoes < 50 )
below_fifty
```
The table displays the year and the film name of all movies with rotten tomatoes scores below 50. The filter function can also be used to filter by categorical variables. For instance, if I wanted just to see the films made in the year 2014, I would code:

```{r filter_cat, exercise=TRUE, exercise.lines = 5}
movieyear = fandango %>%
  select(film, year) %>%
  filter(year == "2014" )
movieyear
```
### Mutate 

*Mutate function can be used to create a new column for your data set. For example, there are a lot of ratings within the data set, so say I wanted to combine as many of the ratings as I could (provided they are all out of the same scale) and see the average rating of the movie. I would:*
Run Code
```{r mutate, exercise=TRUE, exercise.lines = 6}
average_rating = fandango %>%
  mutate(overall_rating = ((rt_norm + rt_user_norm + metacritic_norm + metacritic_user_nom + imdb_norm)/5)) %>%
  select(film, overall_rating)

average_rating
```
This produces an average rating out of a scale of 5 for each film.

### Arrange and Join 
*If we then wanted to view the data to see what top 5 movies have the best rating. We would arrange the data in descending order, but first, we would like to combine the overall average data to the original data set and then view the top 5. there are various types of joins INNER JOIN gets all records that are common between both tables based on the supplied. LEFT JOIN gets all records from the LEFT linked and the related record from the right table, but if you have selected some columns from the RIGHT table, if there are no related records, these columns will contain NULL. RIGHT JOIN is like the above but gets all records in the RIGHT table. FULL JOIN gets all records from both tables and puts NULL in the columns where related records do not exist in the opposite table. For this exercise, we want to use left join to join the average rating table to the fandango table* 

*When using the arrange function alone, it will automatically arrange it in ascending order based on the parameters you give in the parenthesis. To arrange in descending order, You would add on the following code from the previous section. *

Run Code
```{r joinnarrange, exercise=TRUE, exercise.lines = 14}

average_rating = fandango %>%
  mutate(overall_rating = ((rt_norm + rt_user_norm + metacritic_norm + metacritic_user_nom + imdb_norm)/5)) %>%
  select(film, overall_rating)

Top_moviess = fandango %>%
  left_join(average_rating, by = "film")

Tip_movies = Top_moviess %>%
  arrange(desc(overall_rating)) %>%
  select(film, overall_rating) %>%
  top_n(5, wt = overall_rating)

Tip_movies

```
Displays the top 5 movies with respects to their overall average rating.

### Summerize and Group by 
Another step in data wrangling is the summarize and group by function. summarize() and group_by() allow us to make calculations and comparisons. summarize() (with or without group_by()) collapses the data into rows based on summary statistics.The important step in summarize() is specifying how we want to reduce the data to a single set of summary statistics (for example, by mean, median, max, IQR ). 
*Let's say you wanted to group the movies by year to see which movies had better ratings, but to keep in nonbias, you also want to include the number of votes that year got total for each movie rating just for comparison.* 
Run Code
```{r summerize, exercise=TRUE, exercise.lines = 7}
twoyears = fandango %>%
  group_by(year) %>%
  summarise(imdb_rating = mean(imdb), Total_voters = sum(imdb_user_vote_count)) %>%
  select(year, Total_voters, imdb_rating)
twoyears
```
The IMDB rating is on a scale of 0-10. And while the year 2015 had four times as many voters, they still had a lower IMDB rating. Something that can help explain this is the response bias that is an issue involved in individuals filling out surveys. Those who experienced a bad time are more likely to leave an awful review than that of a person who enjoyed their time. 

## Bringing it All Together

### Exercise with Hint

*Here's an exercise where we walk through the entire data wrangling process from start to finish using as many functions as the data needs.*

We wanted to see if the highest voted films and if that affects the difference between the two Fandango ratings. To do this:
We used the mutate function to find the percent difference between the fandango stars rating and the fandango score.
Because there are 23 variables, and most aren't needed for this, we selected only the film name, the number of stars the film had on its Fandango movie page, the Fandango rating value for the film, The number of user votes the film had on Fandango, and the new column, the percent difference. 
We then used the arrange function to order the dataset in descending values from most voted movies. 
Finally, we used the top_n function to only see the top 10 most voted films.
After this, we created a scatter plot to see a visualization of distribution. 

```{r full_excercise, exercise=TRUE, exercise.lines = 10}
data_wrangling_fandango= fandango %>% 
    mutate(percent_difference= ((fandango_stars-fandango_ratingvalue)/2)*100) %>% 
    select(film, fandango_stars, fandango_ratingvalue, fandango_votes, percent_difference) %>%
    arrange(desc(fandango_votes)) %>% 
    top_n(10, wt = fandango_votes) 
data_wrangling_fandango

ggplot(data = data_wrangling_fandango, aes(x = fandango_votes, y = percent_difference)) + 
  geom_point()+
   labs(x = "The Number of Votes the Film Had", y = "Percent Difference (%)", 
       title = "Distribution of Percent Difference Between the Fandango Rating and Fandango Star", 
      subtitle = "in the top ten most voted films")+
  scale_y_continuous(breaks = seq(0, 50, by = 5),
                     labels = c("0", "5", "10", "15", "20", "25", "30", "35", "40", "45", "50"))
```
We noticed that all top 10 highest voted movies had a low difference, three of the movies not having a difference at all. The highest percent difference was 25%, which is still a very low difference. In the graph it looks like that the more percent difference is with a lower amount of votes, which is unexpected. We assumed in the beginning that the more votes the bigger the difference would be. We can conclude that the amount of votes does not affect the percent difference between the two rating styles of Fandango. 

### Quiz

*A short vocab quiz to help distinguish when to use join, inner join, and left join, as it can be confusing at times and I constanly am worrying if I chose the righ join for my data*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which Join Would not return values if the vlalues in table 1 do not match the Values in table 2?",
    answer("Left Join"),
    answer("Right Join"),
    answer("Inner Join", correct = TRUE),
    answer("Full Join")
  ),
  question("Which Join(s) return all rows in Table1, and all columns from Table1 and Table2. Rows in Table1 with no match in Table2 will have NA values in the new columns?",
    answer("Left Join", correct = TRUE),
    answer("Inner Join"),
    answer("Full Join", correct = TRUE),
    answer("Right Join", correct = TRUE), allow_retry = TRUE
  )
)
```

