---
title: "proj 2"
output: html_document
---

Data wrangeling is the process of transforming and mapping data from one "raw" data form into another format with the intent of making it more appropriate and valuable for a variety of downstream purposes such as analytics. Data wrangeling is very useful in coding especially when dealing with a large data set. For the purpose of displaying how one uses R functions to wrangle data we will be using the Fandango data set from the fivethirtyeight package. The Fandango data set is a data frame with 146 rows representing movies and 23 variables. These 23 variables range from the film name, year, and varies ratings. The file will take the reader on a step by step tutorial on how to data wrangle. 

```{r}
# packages that will be needed
library(tidyverse) # Automatically loads ggplot2 and dplyr
library(fivethirtyeight)
```

dplyr is a good package for data wrangling. It is simple for helping to understand and work through manupulating data. One of the greatest advantages of this package is the pipe function. You can use the pipe function “%>%” to write a pipeline combine different functions in R. You can even save "piped" operations as a new data frame. 

Personally I know it can be confusing all of the words used with the pip operation in dplyr so to make things easier dplyr provides a grammar for data transformation.
Five verbs for working with data in a data frame
- select(): takes (selects) a subset of columns (variables)
- filter(): take a subset of rows (observations, cases, individuals), a way to filter the data to only see what you are looking for. 
- mutate(): add or modify existing columns (can even use this function to compute new variables through a provided equation)
- arrange(): sort rows (descending or ascending order, etc.)
- summarize(): aggregate data across rows (e.g., grouped means)
```{r}
view(fandango)
```
first we will try to use the select funtion say you just wanted to know the name of the movie and the year of the film, and its rotten tomato score. you would write the pipline as  
```{r}
movie_year = fandango %>%
  select(film, year, rottentomatoes)
movie_year
```
it displays all 146 movies the year or the film and the rotten tomatoe score, pretty cool right? but 146 rows is still a lot of data.

Next the filter function allows you to take teh data that you selected and filter it to take a subset of rows to only see what you are looking for. Filter can be used alone or in conjunction with the select function. 
```{r}
below_fifty = fandango %>%
  select(film, year, rottentomatoes) %>%
  filter(rottentomatoes < 50 )
below_fifty
```
The table displays the year and the film name of all movies with rotten tomatoe scores below 50. Filter can be used also to filter by categoricol variables. For instance if I wanted to just see the films made in year 2014 I would code
```{r}
movieyear = fandango %>%
  select(film, year) %>%
  filter(year == "2014" )
movieyear
```
Mutate function can be used to create a new column for you data set. For exsample there are alot of ratings within the data set so say I wanted to combine as many of the ratings as I could (provded they are all out of the same scale) and see the average rating of the movie. I would 
```{r}
average_rating = fandango %>%
  mutate(overall_rating = ((rt_norm + rt_user_norm + metacritic_norm + metacritic_user_nom + imdb_norm)/5)) %>%
  select(film, overall_rating)

average_rating
```
Ths produces an avage rating out of a scale of 5 for each film. 

If we then wanted to view the data to see what top 5 movies have the best rating. We would arrange the data in descedning order and view top 5. When using the arrange function alone it will automatically arrange it in ascedning order based off of the parameters you give in the perathesis. To arrange in descending order
```{r}
Top_movies = average_rating %>%
  arrange(desc(overall_rating)) %>%
  top_n(5, wt = overall_rating)

Top_movies
```
Displays the top 5 movies with respects to their overall avaerage rating. 

Another step in data wrangeling is the summerize and group by function. summarize() and group_by() allow us to make calculations and comparisons. summarize() (with or without group_by()) collapses the data into rows based on summary statistics.The important step in summarize() is specifying how we want to reduce the data to a single set of summary statistics (for exsample by mean, median, max, IQR )
Lets say you wanted to group the movies by year to see which movies had better ratings, but in order to keep in non bias you also want to include the amount of votes that year got total for each movie rating just for compairon. You would
```{r}
twoyears = fandango %>%
  group_by(year) %>%
  summarise(imdb_rating = mean(imdb), Total_voters = sum(imdb_user_vote_count)) %>%
  select(year, Total_voters, imdb_rating)


twoyears
```
The imdb rating is on a scale of 0-10. and while the year 2015 had 4 times as many voters they still had a lower imdb rating. Something that can help explain this is the response bias that is an issue involved in indiviuals filling out surveys. those who experiencd a bad time are more likely to leave an awful reviw than that of a person who enjoyed their time. 
